Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-08-22 Masterarbeit_Robotik_Pamuk.tex
3.2 Real-Time Performance Tuning, 3.2.1 BIOS Configurations, 3.2.2 Kernel Configurations, 3.2.3 Host OS Configurations, 3.2.3.1 CPU affinity and Isolation

\subsection{Salamander 4 Bare Metal}\label{subsec:salamander4-bare-metal}
As a reference point, the \texttt{latency} program was executed on Salamander 4 bare metal for a duration of 10 minutes. The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of minimum, average and maximum latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}. Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the variation in latency over the course of said time.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
	\caption[Variation in latency of Salamander 4 bare metal]{Variation in latency of Salamander 4 bare metal}
	\label{fig:gnuplot_max_latency_hardware}
\end{figure}

@@

\subsection{Salamander 4 Virtualization}\label{subsec:salamander4-virtualization}
In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system. Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, Figure~\ref{fig:gnuplot_max_latency_default} depicts the variation in latency.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
	\caption[Variation in latency with default settings]{Variation in latency with default settings}
	\label{fig:gnuplot_max_latency_default}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/01combined/gnuplot_combined_max_latency.png}
	\caption[Initial Comparison of Variation in Latency between Hardware and Virtualization]{Initial Comparison of Variation in Latency between Hardware and Virtualization}
	\label{fig:gnuplot_max_latency_combined}
\end{figure}

@@

In this chapter, the significant initial gap in latency statistics between the virtualized system and the bare metal system is tackled. For this reason, an extensive tuning process is carried out. This involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself. The individual configurations are discussed in detail and the modifications are justified with a clear explanation. The goal is to bring the latency of the virtualized system closer to that of the bare metal system, thereby ensuring deterministic behavior under real-time constraints.

\bigskip \noindent It is important to mention that a real-time system implies determinism at every component of the system ranging from hardware over the kernel to the software and the application on it. As already established earlier, the guest OS Salamamder 4 runs Xenomai and is hence real-time capable. The host system also needs to be aware of the real-time determinism in order to achieve highest possible reliability. The very step of achieving the goal of reducing latency is to apply the PREEMPT-RT patch to the host system. This patch is a set of modifications to the Linux kernel with the goal of making it fully preemptible. This means, almost all parts of the kernel can be preempted and higher-priority tasks can interrupt lower-priority ones. This significantly reduces the time it takes for high-priority tasks to start executing after an event~\cite{RealtimeKernelPatchset}. The patch also includes support for high-resolution timers and makes the system more predictable. Determinism and predictability are key for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch}.

\bigskip \noindent Nevertheless, a real-time kernel alone does not make a system truly “real-time”~\cite{WhatRealtimeLinux}. Additional modifications are required to achieve this. The next sections will deal with these real-time performance tunings. After each tuning, the \texttt{latency} test of Xenomai is executed to evaluate the improved latency. The most important metric of a latency test is without a doubt the worst latency value, because the system is only as reliable as its worst measurement. Since the period of the tests is set to 100\textmu s, there are approximately 100000 samples per second.

@@

BIOS stands for Basic Input/Output System. It abstracts the hardware and enables basic functions of a computer during the booting process, such as starting the operating system and loading other software. Since the BIOS is embedded very deep, its configuration can significantly influence the real-time performance of the system. Table~\ref{tab:bios_configuration} illustrates the specific BIOS settings that have been adjusted for the purpose of real-time performance.

\vspace{2em}
\begin{table}[H]
	\centering
	\caption{BIOS Configurations}
	\label{tab:bios_configuration}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
		\begin{tabular}{|l|l|}
			\hline
			\textbf{Option}               & \textbf{Status} \\
			\hline
			Hyper Threading               & Disabled        \\
			\hline
			Intel SpeedStep®              & Disabled        \\
			\hline
			Intel® Speed Shift Technology & Disabled        \\
			\hline
			C States                      & Disabled        \\
			\hline
			VT-d                          & Enabled         \\
			\hline
		\end{tabular}}
\end{table}

@@

\noindent After these first real-time BIOS tunings, the \texttt{latency} test was run to see, whether this had any significant impact on the determinism of the system. \textcolor{red}{Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS settings.}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
	\caption[Latency taskset]{Latency taskset}
	\label{fig:gnuplot_max_latency_taskset}
\end{figure}

\noindent The maximum latency value recorded was 457.545\textmu s. This is a step for the better compared to the initial version, but certainly still not acceptable within this work's definition of hard real-time determinism.

\subsection{Kernel Configurations}\label{subsec:kernel_configurations}

The kernel command-line parameters are shown in Code~\ref{script:kernel_cli} below.

@@

	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a DMA (Direct Memory Access)-capable I/O bus, such as graphics cards and network adapters, to the main memory. It can enhance device performance by allowing these devices to directly access and use memory, which is especially helpful when these devices are virtualized.
	\item \clearpage
	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel’s RDT (Resource Director Technology). Unlike L1 and L2 caches, where each core has its fixed capacity, L3 cache is a shared pool among multiple cores. L3 CAT is a mechanism that controls the amount of L3 cache that a process can use. By controlling cache allocation, it can prevent a single process from monopolizing the L3 cache, which is particularly beneficial in virtualized environments, where multiple virtual machines share the same physical host.

@@

	\item \textbf{irqaffinity=0}: Sets the default Interrupt Request affinity to none. This means that no CPU core is preferred over another for handling IRQs. Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs. Subsection~\ref{subsubsec:irq_handling} dives deeper into Interrupt Request affinity.

@@

\textcolor{red}{\noindent After the BIOS and Kernel configurations, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} shows }

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/gnuplot_max_latency_rt_kernelparam.png}
	\caption[Latency rt kernelparam]{Latency rt kernelparam}
	\label{fig:gnuplot_max_latency_rt_kernelparam}
\end{figure}

\subsection{Host OS Configurations}\label{subsec:host_configurations}
The host OS needs to provide an environment where the guest OS can operate in real-time. This entails a number of host-side adjustments to reduce interruptions and reduce latency. In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest OS is provided.
\subsubsection{CPU affinity and Isolation}\label{subsubsec:cpu_isolation}

Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved. For CPU isolation, the~\texttt{isolcpus} function was used to isolate a performance CPU from the general scheduling algorithms of the operating system. This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks. However, the isolcpus function only isolates at the user level and does not affect kernel tasks. Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}, including systemd services. To prevent systemd services from running on an isolated CPU, the CPU affinity can be set the with line \texttt{CPUAffinity=0 1 2 3 5 6 7 8 9 10 11 12 13} in the \texttt{/etc/systemd/system.conf} file to indicate the CPUs that systemd-services are allowed to run on. Every CPU other than the isolated CPU 4 is allowed. Code~\ref{output:pse} shows the user and kernel tasks that run on CPU 4. After the isolation, user tasks other than the QEMU process have been removed from running on this CPU. Only few per-CPU kernel threads that are tied to this CPU still take CPU time.

@@

\subsubsection{KVM Entry and KVM Exit}\label{subsubsec:kvm_entry_exit}
\textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host. However, the isolcpus function only isolates CPUs from the scheduler, meaning that user-space processes won't be scheduled on these CPUs, but kernel tasks and interrupts can still utilize the CPU.This led to the investigation of the causes for the observed high and inconsistent latency. The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host. Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host. This frequent switching hinders the guest's ability to run continuously, thereby increasing the virtualization latency.}

\subsubsection{Interrupt Affinity}\label{subsubsec:irq_handling}
Once the CPUs were isolated, interrupt requests handling was the next step. The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another. This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems. The~\texttt{/proc/interrupts} file can be monitored using the command~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time. The IRQs needed to be removed from the isolated CPU by manipulating the~\texttt{/proc/irq/<IRQ>/smp\_affinity} files. The value in the~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU13). When there are 14 CPUs available, the default value for~\texttt{smp\_affinity} would be 3FFF. Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 3FEF (11 1111 1110 1111). The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU. By changing the values of the~\texttt{smp\_affinity} files, the assignment of IRQs to CPUs is controlled.

\clearpage
\vspace*{-5em}
\begin{lstlisting}[name={Check distribution of interrupt requests across each CPU},label={script:smp_affinity}]
		import os
		import pandas as pd
		from tabulate import tabulate

@@

\noindent However, as the proc filesystem resets to its default state after each reboot, manually changing numerous IRQ files is tedious and time-consuming. This process was automated using the shell script in Code~\ref{script:change_smp_affinity}, which is executed after every reboot. Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads. This can be done through finding out the respective process IDs through \texttt{ps -e | grep irq/.*nvme} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.

@@

>>>>

\subsection{Salamander 4 Bare Metal}\label{subsec:salamander4-bare-metal}
As a reference point, the \texttt{latency} program was executed on Salamander 4 bare metal for a duration of 10 minutes. The complete command used was~\texttt{latency -h -g gnuplot.txt -T 600}, which runs the latency measurement tool for 600 seconds and prints histograms of minimum, average and maximum latencies in a Gnuplot-compatible format to the file~\texttt{gnuplot.txt}. Figure~\ref{fig:gnuplot_max_latency_hardware} depicts the distribution of latency over the course of said time.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/0hardware/gnuplot_max_latency_hardware.png}
	\caption[Distribution of Latency of Salamander 4 Bare Metal]{Distribution of Latency of Salamander 4 Bare Metal}
	\label{fig:gnuplot_max_latency_hardware}
\end{figure}

@@

\subsection{Salamander 4 Virtualization}\label{subsec:salamander4-virtualization}
In addition to providing Salamander 4 on its own hardware, SIGMATEK has also developed a virtualised version of this operating system. Measuring the latency of the Salamander 4 virtualization with the default QEMU script in Code~\ref{script:qemu_def} and no further adjustments for 10 minutes, Figure~\ref{fig:gnuplot_max_latency_default} depicts the distribution of latency.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/1default/gnuplot_max_latency_default.png}
	\caption[Distribution of Latency of Salamander 4 unmodified Virtualization]{Distribution of Latency of Salamander 4 unmodified Virtualization}
	\label{fig:gnuplot_max_latency_default}
\end{figure}

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/01combined/gnuplot_combined_max_latency.png}
	\caption[Initial Comparison of Latency Distribution between Hardware and Virtualization]{Initial Comparison of Latency Distribution between Hardware and Virtualization}
	\label{fig:gnuplot_max_latency_combined}
\end{figure}

@@

In this chapter, the significant initial gap in latency statistics between the virtualized system and the bare metal system is tackled. For this reason, an extensive tuning process is carried out. This involves configurations spanning the BIOS, kernel, host OS, QEMU/KVM virtualization layer, and the Salamander 4 OS itself. The individual configurations are discussed in detail and the modifications are justified with a clear explanation. The goal is to bring the latency of the virtualization closer to that of the bare metal, thereby ensuring deterministic behavior under real-time constraints.

\bigskip \noindent It is important to mention that a real-time system implies determinism at every component of the system, ranging from hardware over the kernel to the software and the application on it. As already established earlier, the guest OS Salamamder 4 runs Xenomai and is hence real-time capable. The host system also needs to be aware of the real-time determinism in order to achieve highest possible reliability. The very step of achieving the goal of reducing latency is to apply the PREEMPT-RT patch to the host system. This patch is a set of modifications to the Linux kernel with the goal of making it fully preemptible. This means, almost all parts of the kernel can be preempted and higher-priority tasks can interrupt lower-priority ones. This significantly reduces the time it takes for high-priority tasks to start executing after an event~\cite{RealtimeKernelPatchset}. The patch also includes support for high-resolution timers and makes the system more predictable. Determinism and predictability are key for real-time applications where the timing of operations must be guaranteed with precise timing and scheduling~\cite{rostedtInternalsRTPatch}.

\bigskip \noindent Nevertheless, a real-time kernel alone does not make a system truly “real-time”~\cite{WhatRealtimeLinux}. Additional modifications are required to achieve this. The next sections will deal with these real-time performance tunings. After each tuning, the \texttt{latency} test of Xenomai is executed to evaluate the improved latency. The most important metric of a latency test is without a doubt the worst latency value, because the system is only as reliable as its worst measurement. Since the period of the tests is set to 100\textmu s, there are approximately 1,000,000 samples per second.

@@

BIOS stands for Basic Input/Output System. It abstracts the hardware and enables basic functions of a computer during the booting process, such as starting the operating system and loading other software. Since the BIOS is embedded very deep, its configuration can significantly influence the real-time performance of the system. Table~\ref{tab:bios_configuration} illustrates the specific BIOS settings that have been adjusted for the purpose of real-time performance. As an important note, this is not a definitive, full-ended list, and other devices may have additional settings that can be modified to achieve even better latency.

\begin{table}[H]
	\centering
	\caption{BIOS Configurations}
	\label{tab:bios_configuration}
	\setlength{\tabcolsep}{0.5em} % for the horizontal padding
	{\renewcommand{\arraystretch}{1.2}% for the vertical padding
		\begin{tabular}{|l|l|}
			\hline
			\textbf{Option}               & \textbf{Status} \\
			\hline
			Hyper Threading               & Disabled        \\
			\hline
			Intel SpeedStep®              & Disabled        \\
			\hline
			Intel® Speed Shift Technology & Disabled        \\
			\hline
			C States                      & Disabled        \\
			\hline
			VT-d                          & Enabled         \\
			\hline
		\end{tabular}}
\end{table}

@@

\noindent After these first real-time tunings in the BIOS settings, the \texttt{latency} test was run to see, whether this had any significant impact on the determinism of the system. Figure~\ref{fig:gnuplot_max_latency_taskset} shows the latency values after tuning BIOS settings.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/2taskset/gnuplot_max_latency_taskset.png}
	\caption[Latency Distribution after tuning BIOS Settings]{Latency Distribution after tuning BIOS Settings}
	\label{fig:gnuplot_max_latency_taskset}
\end{figure}

\noindent The maximum latency value recorded was 457.545\textmu s. When compared to the inital worst latency of 707.62 \textmu s, this is a step for the better, but certainly still not acceptable within this work's definition of hard real-time determinism.

\subsection{Kernel Configurations}\label{subsec:kernel_configurations}
The kernel command-line parameters are shown in Code~\ref{script:kernel_cli} below. To access and modify them, the file is located in \texttt{/etc/default/grub}.

@@

	\item \textbf{intel\_iommu=on}: Enables Intel's IOMMU (Input/Output Memory Management Unit), which connects a DMA (Direct Memory Access)-capable I/O bus, such as graphics cards and network adapters, to the main memory. It allows these devices to directly access and use memory, which is especially helpful when they are virtualized.
	\item \textbf{rdt=l3cat}: Activates the L3 CAT (L3 Cache Allocation Technology) feature of Intel’s RDT (Resource Director Technology). Unlike L1 and L2 caches, where each core has its fixed capacity, L3 cache is a shared pool among multiple cores. L3 CAT controls the amount of L3 cache that a process can use. By controlling cache allocation, it can prevent a single process from monopolizing the L3 cache, which is particularly beneficial in virtualized environments, where multiple virtual machines share the same physical host.

@@

	\item \textbf{irqaffinity=0}: Sets the default interrupt request affinity to none. This means that no CPU core is preferred over another for handling IRQs. Instead, it lets the operating system decide how to distribute these IRQs across all the CPUs. Subsubsection~\ref{subsubsec:irq_handling} dives deeper into interrupt request affinity.

@@

\noindent After the kernel modifications on top of the BIOS settings from the last subsection, Figure~\ref{fig:gnuplot_max_latency_rt_kernelparam} demonstrates the latency
values gathered in 10 minutes.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/4rt_kernelparam/gnuplot_max_latency_rt_kernelparam.png}
	\caption[Latency Distribution after tuning Kernel Settings]{Latency Distribution after tuning Kernel Settings}
	\label{fig:gnuplot_max_latency_rt_kernelparam}
\end{figure}

\noindent Apparent is a significant improvement in the latency of the virtualized Salamander 4 OS. With a worst latency value of 21.694\textmu s, the real-time tunings appear to be working as desired.

\subsection{Host OS Configurations}\label{subsec:host_configurations}
The host OS needs to provide an environment where the guest OS can operate in real-time. This entails a number of host-side adjustments to reduce interruptions and latency. In the following, a detailed overview of these configurations and their impact on the real-time performance of the guest Salamander 4 OS is provided.
\subsubsection{CPU affinity and Isolation}\label{subsubsec:cpu_isolation}

Isolating CPUs means removing all user-space threads and unbound kernel threads, since bound kernel threads are tied to specific CPUs and hence cannot be moved. For CPU isolation, the~\texttt{isolcpus} function is used to isolate a performance CPU from the general scheduling algorithms of the operating system. This means that the isolated CPUs will not be used for regular task scheduling, allowing them to be dedicated for the real-time specific tasks. However, the isolcpus function only isolates at the user level and does not affect kernel tasks. Consequently, these kernel tasks and interrupts can still utilize the CPU~\cite{maPerformanceTuningKVMbased}, including systemd services. To prevent systemd services from running on an isolated CPU, the CPU affinity can be set with the line \texttt{CPUAffinity=0 1 2 3 5} in the \texttt{/etc/systemd/system.conf} file to indicate the CPUs that systemd services are allowed to run on. Every CPU other than the isolated CPU 4 is allowed. Code~\ref{output:pse} shows the user and kernel tasks that run on CPU 4. An isolated CPU can be bound to a task by using \texttt{taskset -c <CPU>}, as shown later in Subsection~\ref{subsec:guest_configurations}. After the isolation, user tasks other than the QEMU process have been removed from running on this CPU. Only few per-CPU kernel threads that are tied to this CPU still take CPU time.

@@

\subsubsection{Interrupt Affinity}\label{subsubsec:irq_handling}
Once the CPUs were isolated, interrupt requests handling was the next step. The purpose of interrupt requests is to inform the CPU to stop working on a certain job and start working on another. This allows hardware devices to communicate with the CPU through frequent context switches, which can introduce latency in real-time systems. All the existing interrupts can be monitored using the command~\texttt{watch -d -n 1 cat /proc/interrupts} to observe changes in the interrupt requests handled by each CPU in real-time. The IRQs needed to be removed from the isolated CPU by manipulating the~\texttt{/proc/irq/<IRQ>/smp\_affinity} files. The value in the~\texttt{smp\_affinity} file is a bitmask in hexadecimal format where each bit corresponds to a CPU. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the most significant bit (MSB) on the left corresponds to the last CPU (CPU5). When there are 6 CPUs available, the default value for~\texttt{smp\_affinity} would be 3F (11 1111). Removing CPU 4 out of this bitmask would be setting bit five to zero, resulting in 2F (10 1111). The python script in Code~\ref{script:smp_affinity} was written to show a table of the distribution of interrupt requests across each CPU, since the \texttt{/proc/interrupts} file does not give away this information and looking individually in the \texttt{/proc/irq/<IRQ>/smp\_affinity} files for every IRQ would be time-consuming. The program first reads the CPU affinity settings for each IRQ from the \texttt{/proc/irq} directory and stores the CPUs assigned to each IRQ in a dictionary. It then creates a data frame to represent this information in a table format, marking the assigned CPUs with ‘x’. Finally, it prints the table and saves it as a markdown file. Figure~\ref{fig:interrupt_affinity_table} shows the output of the program.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\columnwidth]{masterthesis-documentation/docs/sigmatek/salamander4/latency_reduction/IRQaffinity/table_CPU_IRQ.png}
	\caption[Interrupt Affinity Table]{Interrupt Affinity Table}
	\label{fig:interrupt_affinity_table}
\end{figure}
\clearpage
\vspace*{-5em}
\begin{lstlisting}[name={Check distribution of interrupt requests across each CPU},label={script:smp_affinity}]
		import os
		import pandas as pd
		from tabulate import tabulate

@@

\noindent However, as the \texttt{proc} filesystem resets to its default state after each reboot, manually changing numerous IRQ files through the said bitbasks is tedious and time-consuming. This process can be was automated using the shell script in Code~\ref{script:change_smp_affinity}, which can automatically be executed after every reboot.

Additionally, the CPU affinity of IRQ threads of NVME (Non-Volatile Memory Express, a type of SSD storage) can be set away from CPU 4 to avoid impacting real-time workloads. This can be done through finding out the respective process IDs through \texttt{ps -e | grep irq/.*nvme} and then executing \texttt{sudo taskset -a -p -c 0 <PID>}.

@@

\section{KVM Entry and KVM Exit}\label{subsubsec:kvm_entry_exit}
\textcolor{red}{\noindent Upon isolating a CPU to the QEMU process, it was anticipated that the guest would utilize nearly 100\% of the CPU's capacity, with minimal to no intervention from the host. However, the isolcpus function only isolates CPUs from the scheduler, meaning that user-space processes won't be scheduled on these CPUs, but kernel tasks and interrupts can still utilize the CPU.This led to the investigation of the causes for the observed high and inconsistent latency. The guest operates within the~\texttt{kvm\_entry} and~\texttt{kvm\_exit} events of the host. Kernelshark revealed a high frequency of~\texttt{kvm\_exit} events, indicating that the guest frequently relinquishes control of the CPU back to the host. This frequent switching hinders the guest's ability to run continuously, thereby increasing the virtualization latency.}
